---
title: "eMLR model loop"
author: "Donghe Zhu and Dr Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```


```{r library, message = FALSE, warning = FALSE}
library(tidyverse)
library(patchwork)
library(stars)
library(collapse)
library(lubridate)
library(rqdatatable)
library(gsw)
```

```{r read_parameters, include = FALSE}
parameters <-
  read_rds(here::here("data",
                       "parameters.rds"))
```

```{r read_mask_files, include = FALSE}

basinmask <- read_csv(
  here::here(
    "data/World_Ocean_Atlas_2018/_summarized_files",
    "basin_mask_WOA18_AIP.csv"
  )
)
basinmask <- basinmask %>%
  select(-basin)

landmask <-  read_csv(
  here::here(
    "data/World_Ocean_Atlas_2018/_summarized_files",
    "land_mask_WOA18.csv"
  )
)

```

```{r read_section_global_coordinates, include = FALSE}

section_global_coordinates <- read_csv(
  here::here(
    "data/World_Ocean_Atlas_2018/_summarized_files",
    "section_global_coordinates.csv"
  )
)

```

```{r read_functions, include = FALSE}
source(here::here("code", "plotting_functions.R"))
```

```{r ggplot_theme, include = FALSE}
# ggplot theme for all plots
theme_set(theme_bw())
```


# Read GLODAP clean data

```{r read_GLODAP_clean_data}

GLODAP <- read_csv(here::here("data/GLODAPv2_2020",
                              "GLODAPv2.2020_Merged_clean.csv"))

```

# Subset model data according to GLODAP observation

Here we subset regridded (1x1) data from the model with variable forcing, according to the presence of GLODAP observations in a previously cleaned file.

Simultaneously, we calculate annual model summary statistics and annual average DIC fields.

Besides, Model results are given in [mmol m^-3^], whereas GLODAP data are in [µmol kg^-1^]. This refers to the variables:

- DIC  
- ALK  
- O2  
- NO3  
- PO4  
- SiO3  

For comparison, model results were converted from [mmol m^-3^] to [µmol kg^-1^]

```{r set_absolute_path}

# set path to model output files
path_model_cmorized <-
  "/net/kryo/work/loher/CESM_output/RECCAP2/cmorized_Dec2020/split_monthly/"

# set path to model analysis/subsetting output files
path_model_analysis <-
  "/net/kryo/work/loher/CESM_output/RECCAP2/analysis/"

```


```{r subset_cmorized_variable_forcing_data, eval=FALSE}

# set name of model to be subsetted
model_ID <- "A"

# for loop across years
years <- c("1982":"2019")

for (i_year in years) {
  # for manual testing
  # i_year <- years[1]
  
  # select GLODAP data for that year
  GLODAP_year <- GLODAP %>%
    filter(year == i_year)
  
  # create month x lat x lon grid of observations
  grid_subset <- GLODAP_year %>%
    select(month, lat, lon) %>%
    unique()
  
  # create month x lat x lon x depth grid of observations
  obs_depth <- GLODAP_year %>%
    select(month, lat, lon, depth) %>%
    unique()
  
  # for loop across variables
  variables <-
    c("so", "thetao", "dissic", "talk", "o2", "no3", "po4", "si")
  # "SALT", "TEMP", "DIC", "ALK", "O2", "NO3", "PO4", "SiO3"
  
  for (i_variable in variables) {
    # Uncomment line below for manual testing
    # i_variable <- variables[1]
    
    # read list of all files
    file <-
      list.files(
        path = path_model_cmorized,
        pattern = paste(
          "^",
          i_variable,
          "_CESM-ETHZ_",
          model_ID,
          "_1_gr_",
          i_year,
          ".nc",
          sep = ""
        )
      )
    print(file)
    
    # read in data
    variable_data <-
      read_ncdf(paste(path_model_cmorized,
                      file,
                      sep = ""))
    
    # convert to tibble
    variable_data_tibble <- variable_data %>%
      as_tibble()
    
    # remove open link to nc file
    rm(variable_data)
    
    # remove na values -> do this early, to speed up subsequent tasks
    variable_data_tibble <-
      variable_data_tibble %>%
      filter(!is.na(!!sym(i_variable)))
    
    # convert longitudes
    variable_data_tibble <- variable_data_tibble %>%
      mutate(lon = if_else(lon < 20, lon + 360, lon))
    
    # mutate variables and filter NaN value
    variable_data_tibble <- variable_data_tibble %>%
      mutate(month = month(time_mon),!!sym(i_variable) := as.numeric(!!sym(i_variable))) %>%
      select(-time_mon)
    
    # calculate density
    if (exists("unit_trans")) {
      unit_trans <- left_join(unit_trans, variable_data_tibble)
    }
    
    if (!exists("unit_trans")) {
      unit_trans <- variable_data_tibble
    }
    
    if (i_variable == "thetao") {
      unit_trans <- unit_trans %>%
        mutate(
          rho = gsw_pot_rho_t_exact(
            SA = so,
            t = thetao,
            p = depth,
            p_ref = 10.1325
          ),
          dens_factor = 1000000 / rho
        ) %>%
        select(-so, -thetao)
    }
    
    # unit transfer from mol/m3 to µmol/kg
    if (i_variable != "so" && i_variable != "thetao") {
      unit_trans <- unit_trans %>%
        mutate(!!sym(i_variable) := !!sym(i_variable) * dens_factor)
      variable_data_tibble <- unit_trans %>%
        select(lon, lat, depth,!!sym(i_variable), month)
      unit_trans <- unit_trans %>%
        select(-!!sym(i_variable))
    }
    
    # calculate model summary stats
    stats <- variable_data_tibble %>%
      pull(!!sym(i_variable)) %>%
      summary()
    
    stats <- c(year = i_year, variable = i_variable, stats)
    
    if (exists("stats_summary")) {
      stats_summary <- bind_rows(stats_summary, stats)
    }
    
    if (!exists("stats_summary")) {
      stats_summary <- stats
    }
    
    rm(stats)
    
    # subset model at month x lat x lon grid of observations
    model_subset <-
      inner_join(grid_subset, variable_data_tibble)
    
    # join model and month x lat x lon x depth grid of observations
    model_obs <- full_join(model_subset, obs_depth)
    
    # calculate nr of observations per month x lat x lon grid
    model_obs <- model_obs %>%
      group_by(month, lat, lon) %>%
      mutate(n = sum(!is.na(!!sym(i_variable)))) %>%
      ungroup()
    
    # interpolate variable value from model to observation depth
    model_obs_interpo <- model_obs %>%
      filter(n > 1) %>%
      group_by(lon, lat, month) %>%
      arrange(depth) %>%
      mutate(!!sym(i_variable) := approxfun(depth,!!sym(i_variable), rule = 2)(depth)) %>%
      ungroup()
    
    # subsetted interpolated values at observation depth
    model_obs_interpo <-
      inner_join(obs_depth, model_obs_interpo) %>%
      select(-n) %>%
      mutate(year = as.numeric(i_year))
    
    # rename interpolated model variable to indicate as model output
    model_obs_interpo <- model_obs_interpo %>%
      rename(!!sym(paste(i_variable, "model", sep = "_")) := !!sym(i_variable))
    
    # add model subset to GLODAP
    GLODAP <-
      natural_join(
        GLODAP,
        model_obs_interpo,
        by = c("year", "month", "lat", "lon", "depth"),
        jointype = "FULL"
      )
    
    # calculate annual average variable
    variable_data_tibble_annual_average <- variable_data_tibble %>%
      fselect(-month) %>%
      fgroup_by(lat, lon, depth) %>% {
        add_vars(fgroup_vars(., "unique"),
                 fmean(., keep.group_vars = FALSE))
      }
    
    # select surface annual average variable
    variable_data_tibble_annual_average_surface <-
      variable_data_tibble_annual_average %>%
      filter(depth == min(depth))
    
    # surface map of variable
    surface_GLODAP(
      variable_data_tibble_annual_average_surface,
      obs_depth,
      model_obs_interpo,
      i_variable
    )
    
    ggsave(
      here::here(
        "output/model_subsetting/surface_cmorized_GLODAP_obs",
        paste(i_variable, "_", i_year, ".png", sep = "")
      ),
      width = 5,
      height = 3
    )
    # GLODAP with model subset:red point
    # GLODAP without model subset:black point
    
    # combine annual average model file together
    if (i_variable == "dissic") {
      # write file for annually averaged model field
      variable_data_tibble_annual_average %>%
        write_csv(paste(
          paste(path_model_analysis, "cmorized_clim/", model_ID, sep = ""),
          paste("/dissic_annual_average_", i_year, ".csv", sep = ""),
          sep = ""
        ))
      
    }
    
  }
  
  rm(unit_trans)
  
}

# write file for GLODAP observations + subsetted model variables
GLODAP %>%
  write_csv(paste(
    paste(path_model_analysis, "glodapv2_cmorized_synthetic/", sep = ""),
    "GLODAP_with_cmorized_subset.csv",
    sep = ""
  ))

# write file for model summary statistics
stats_summary %>%
  write_csv(paste(
    paste(path_model_analysis, "cmorized_diag/", sep = ""),
    paste(model_ID, "summary_stats.csv", sep = "_"),
    sep = ""
  ))

```

# Test for interpolation

```{r test_for_interpolation, eval=FALSE}

# subset model at month x lat x lon grid of observations
model_subset <-
  inner_join(grid_subset, variable_data_tibble)

# join model and month x lat x lon x depth grid of observations
model_obs <- full_join(model_subset, obs_depth)

# calculate nr of observations per month x lat x lon grid
model_obs <- model_obs %>%
  group_by(month, lat, lon) %>%
  mutate(n = sum(!is.na(!!sym(i_variable)))) %>%
  ungroup()

# interpolate variable value from model to observation depth
model_obs_interpo <- model_obs %>%
  filter(n > 1) %>%
  group_by(lon, lat, month) %>%
  arrange(month, lat, lon, depth) %>%
  mutate(!!sym(paste(i_variable, "interpolate", sep = "_")) := approxfun(depth, !!sym(i_variable), rule = 2)(depth)) %>%
  ungroup()

ggplot() +
  geom_path(
    data = model_obs_interpo %>%
      filter(lat == 39.5, lon == 348.5,!is.na(!!sym(i_variable)), month == 11) %>%
      arrange(depth),
    aes(!!sym(i_variable), depth, col = "model")
  ) +
  geom_point(
    data = model_obs_interpo %>%
      filter(lat == 39.5, lon == 348.5, month == 11),
    aes(!!sym(
      paste(i_variable, "interpolate", sep = "_")
    ), depth, col = "interpolated")
  ) +
  geom_point(
    data = model_obs_interpo %>%
      filter(lat == 39.5, lon == 348.5,!is.na(!!sym(i_variable)), month == 11) %>%
      arrange(depth),
    aes(!!sym(i_variable), depth, col = "model")
  ) +
  scale_y_reverse() +
  scale_color_brewer(palette = "Dark2", name = "") +
  labs(title = paste("Interpolation to sampling depth -", i_variable, "of year", i_year))

```

# Annual average DIC fields

To derive Cant estimates for certain years, we calculate annual average DIC fields also for the output of a model with normal forcing.

```{r calculate_normal_forcing_annual_average_DIC, eval=FALSE}

# set name of modelled to be subsetted
model_ID <- "B"

# for loop across years
years <- c("1982":"2019")

for (i_year in years) {
  # for manual testing
  # i_year <- years[1]
  
  # read list of all files
  variables <-
    c("so", "thetao", "dissic")
  
  for (i_variable in variables) {
    #  i_variable <- variables[1]
    
    file <-
      list.files(
        path = path_model_cmorized,
        pattern = paste(
          "^",
          i_variable,
          "_CESM-ETHZ_",
          model_ID,
          "_1_gr_",
          i_year,
          ".nc",
          sep = ""
        )
      )
    print(file)
    
    # read in data
    temp <-
      read_ncdf(paste(path_model_cmorized,
                      file,
                      sep = "")) %>%
      as_tibble()
    
    if (exists("normal_data")) {
      normal_data <- full_join(normal_data, temp)
    }
    
    if (!exists("normal_data")) {
      normal_data <- temp
    }
    
  }
  
  # remove na values -> do this early, to speed up subsequent tasks
  normal_data_tibble <-
    normal_data %>%
    drop_na()
  rm(temp, i_variable, variables, normal_data)
  
  # convert longitudes
  normal_data_tibble <- normal_data_tibble %>%
    mutate(lon = if_else(lon < 20, lon + 360, lon))
  
  # mutate variables and filter NaN value
  normal_data_tibble <- normal_data_tibble %>%
    mutate(
      month = month(time_mon),
      so = as.numeric(so),
      thetao = as.numeric(thetao),
      dissic = as.numeric(dissic)
    ) %>%
    select(-time_mon)
  
  # calculate density and transfer unit from mol/m3 to µmol/kg
  normal_data_tibble <- normal_data_tibble %>%
    mutate(
      rho = gsw_pot_rho_t_exact(
        SA = so,
        t = thetao,
        p = depth,
        p_ref = 10.1325
      ),
      dens_factor = 1000000 / rho,
      dissic = dissic * dens_factor
    ) %>%
    select(-so, -thetao,-rho,-dens_factor)
  
  # calculate annual average DIC
  normal_data_tibble_annual_average <- normal_data_tibble %>%
    fselect(-month) %>%
    fgroup_by(lat, lon, depth) %>% {
      add_vars(fgroup_vars(., "unique"),
               fmean(., keep.group_vars = FALSE))
    }
  
  # write file for annually averaged model DIC
  normal_data_tibble_annual_average %>%
    write_csv(paste(
      paste(path_model_analysis, "cmorized_clim/", model_ID, sep = ""),
      paste("/dissic_annual_average_", i_year, ".csv", sep = ""),
      sep = ""
    ))
  
}

```

# Calculate model Cant

The model Cant can be calculate as the difference between two model runs. Currently, this is performed for year 1994 and 2007.

```{r calculate_Cant}

# read in variable and normal forcing DIC data
years <- c("1994", "2007")

for (i_year in years) {
  # for manual testing
  i_year <- years[1]
  
  # read in variable forcing annual average DIC data
  variable_annual_average_DIC <-
    read_csv(paste(
      paste(path_model_analysis, "cmorized_clim/A/", sep = ""),
      paste("dissic_annual_average_", i_year, ".csv", sep = ""),
      sep = ""
    )) %>%
    rename(DIC_variable = dissic)
  
  # read in normal forcing annual average DIC data
  normal_annual_average_DIC <-
    read_csv(paste(
      paste(path_model_analysis, "cmorized_clim/B/", sep = ""),
      paste("dissic_annual_average_", i_year, ".csv", sep = ""),
      sep = ""
    )) %>%
    rename(DIC_normal = dissic)
  
  # calculate Cant
  cant <-
    full_join(variable_annual_average_DIC, normal_annual_average_DIC) %>%
    mutate(Cant = DIC_variable - DIC_normal)
  
  rm(variable_annual_average_DIC, normal_annual_average_DIC)
  
  cant <- inner_join(basinmask, cant)
  
  # global surface Cant plot
  print(map_global(cant, "Cant"))
  
  # global section Cant plot
  print(section_global(cant, "Cant"))
  
  # zonal section Cant plot
  for (i_basin_AIP in unique(cant$basin_AIP)) {
    breaks <- c(seq(-30, 70, 10), Inf)
    print(section_zonal(cant,
                        i_basin_AIP = i_basin_AIP,
                        "Cant"))
  }
}

```

