---
title: "Model random subsetting"
author: "Donghe Zhu and Dr Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```


```{r library, message = FALSE, warning = FALSE}
library(tidyverse)
library(patchwork)
library(stars)
library(collapse)
library(lubridate)
library(rqdatatable)
library(gsw)
library(marelac)
```

```{r read_parameters, include = FALSE}
parameters <-
  read_rds(here::here("data",
                       "parameters.rds"))
```

```{r read_mask_files, include = FALSE}

basinmask <- read_csv(
  here::here(
    "data/World_Ocean_Atlas_2018/_summarized_files",
    "basin_mask_WOA18_AIP.csv"
  )
)
basinmask <- basinmask %>%
  select(-basin)

landmask <-  read_csv(
  here::here(
    "data/World_Ocean_Atlas_2018/_summarized_files",
    "land_mask_WOA18.csv"
  )
)

```

```{r read_section_global_coordinates, include = FALSE}

section_global_coordinates <- read_csv(
  here::here(
    "data/World_Ocean_Atlas_2018/_summarized_files",
    "section_global_coordinates.csv"
  )
)

```

```{r read_functions, include = FALSE}
source(here::here("code", "plotting_functions.R"))
```

```{r ggplot_theme, include = FALSE}
# ggplot theme for all plots
theme_set(theme_bw())
```


# Read GLODAP clean data and apply the basinmask

```{r read_GLODAP_clean_data}

GLODAP <- read_csv(here::here("data/GLODAPv2_2020",
                              "GLODAPv2.2020_Merged_clean.csv"))
GLODAP <- inner_join(GLODAP, basinmask) %>%
  select(-basin_AIP)

```

# Randomly subset cmorized model data

Here we randomly subset cmorized (1x1) data from the model with variable forcing, according to the total number of GLODAP observations for the whole period from a previously cleaned file. The number for the annual subset remains the same for each year, which could be expressed by the total number of observations divided by number of years.

Besides, Model results are given in [mol m^-3^], whereas GLODAP data are in [µmol kg^-1^]. This refers to the variables:

- DIC  
- ALK  
- O2  
- NO3  
- PO4  
- SiO3  
- AOU (calculated)

For comparison, model results were converted from [mol m^-3^] to [µmol kg^-1^]

```{r set_absolute_path}

# set path to model output files
path_model_cmorized <-
  "/nfs/kryo/work/loher/CESM_output/RECCAP2/cmorized_Dec2020/split_monthly/"

# set path to model analysis/subsetting output files
path_model_analysis <-
  "/nfs/kryo/work/loher/CESM_output/RECCAP2/analysis/"

```


```{r random_sample_cmorized_variable_forcing_data, eval=FALSE}

# read in a random model
model <-
  read_ncdf(
    "/net/kryo/work/loher/CESM_output/RECCAP2/cmorized_Dec2020/split_monthly/dissic_CESM-ETHZ_A_1_gr_1982.nc"
  ) %>%
  as_tibble() %>%
  drop_na()

# convert longitudes and mutate month
model <- model %>%
  mutate(lon = if_else(lon < 20, lon + 360, lon)) %>%
  mutate(month = month(time_mon))

# only consider model grids within basinmask
model <- inner_join(model, basinmask)

# model grid with depth
model_grid_depth <- model %>%
  select(month, lat, lon, depth)

rm(model)

# set name of model to be subsetted
model_ID <- "A"

# for loop across years
years <- c("1982":"2019")

# set equal number of random model sampling will be made for each year
n = floor(nrow(GLODAP) / length(years))

for (i_year in years) {
  # for manual testing
 # i_year <- years[1]
  
  # random sample n = GLODAP / years from model grid depth
  model_resample_grid_depth <- sample_n(model_grid_depth, n) %>%
    arrange(lat, lon, depth, month) %>%
    mutate(year = i_year)
  
  # for loop across variables
  variables <-
    c("so", "thetao", "dissic", "talk", "o2", "no3", "po4", "si")
  # "SALT", "TEMP", "DIC", "ALK", "O2", "NO3", "PO4", "SiO3"
  
  for (i_variable in variables) {
    # Uncomment line below for manual testing
    #i_variable <- variables[1]
    
    # read list of all files
    file <-
      list.files(
        path = path_model_cmorized,
        pattern = paste(
          "^",
          i_variable,
          "_CESM-ETHZ_",
          model_ID,
          "_1_gr_",
          i_year,
          ".nc",
          sep = ""
        )
      )
    print(file)
    
    # read in data
    variable_data <-
      read_ncdf(paste(path_model_cmorized,
                      file,
                      sep = ""))
    
    # convert to tibble
    variable_data_tibble <- variable_data %>%
      as_tibble()
    
    # remove open link to nc file
    rm(variable_data)
    
    # remove na values -> do this early, to speed up subsequent tasks
    variable_data_tibble <-
      variable_data_tibble %>%
      filter(!is.na(!!sym(i_variable)))
    
    # convert longitudes
    variable_data_tibble <- variable_data_tibble %>%
      mutate(lon = if_else(lon < 20, lon + 360, lon))
    
    # only consider model grids within basinmask
    variable_data_tibble <- inner_join(variable_data_tibble, basinmask) %>%
      select(-basin_AIP)
    
    # mutate variables
    variable_data_tibble <- variable_data_tibble %>%
      mutate(month = month(time_mon), !!sym(i_variable) := as.numeric(!!sym(i_variable))) %>%
      select(-time_mon)
    
    # random sample for each model variable in specific year
    model_resample_grid_depth <-
      left_join(model_resample_grid_depth, variable_data_tibble)
  }
  
  # add random sample model subset for each year together
  if (exists("model_resample")) {
    model_resample <-
      bind_rows(model_resample, model_resample_grid_depth)
  }
  
  if (!exists("model_resample")) {
    model_resample <- model_resample_grid_depth
  }
  
}

# unit transfer from mol/m3 to µmol/kg
model_resample <- model_resample %>%
  mutate(
    rho = gsw_pot_rho_t_exact(
      SA = so,
      t = thetao,
      p = depth,
      p_ref = 10.1325
    ),
    dissic = dissic * (1000000 / rho),
    talk = talk * (1000000 / rho),
    o2 = o2 * (1000000 / rho),
    no3 = no3 * (1000000 / rho),
    po4 = po4 * (1000000 / rho),
    si = si * (1000000 / rho)
  )

# calculate AOU
model_resample <- model_resample %>%
  mutate(
    oxygen_sat_m3 = gas_satconc(
      S = so,
      t = thetao,
      P = 1.013253,
      species = "O2"
    ),
    oxygen_sat_kg = oxygen_sat_m3 * (1000 / rho),
    aou = oxygen_sat_kg - o2
  ) %>%
  select(-oxygen_sat_kg,-oxygen_sat_m3)

# write file for random model sampling
model_resample %>%
  write_csv(paste(
    paste(path_model_analysis, "glodapv2_cmorized_synthetic/", sep = ""),
    "random_subset.csv",
    sep = ""
  ))

```


```{r plot_random_sample_cmorized_data_grid}

# read in random model sampling file
model_resample <-
  read_csv(paste(
    paste(path_model_analysis, "glodapv2_cmorized_synthetic/", sep = ""),
    "random_subset.csv",
    sep = ""
  ))

# plot random sampling cmorized grids in each year
years <- c("1982", "1990", "2000", "2010", "2019")
for (i_year in years) {
  # i_year <- years[1]
  
  model_resample_year <- model_resample %>%
    filter(year == i_year)
  
  print(
    ggplot() +
      geom_bin2d(data = model_resample_year,
                 aes(lon, lat),
                 binwidth = 1) +
      scale_fill_viridis_c(direction = -1) +
      coord_quickmap(expand = 0) +
      labs(
        title = paste("Random Model Sampling of year", i_year),
        subtitle = paste("Nr of observations", nrow(model_resample_year)),
        x = "Longitude",
        y = "Latitude"
      ) +
      geom_raster(data = landmask,
                  aes(lon, lat), fill = "grey80")
  )

}

# Calculate and plot depth distribution of model subset
model_resample_depth <- model_resample %>%
  count(depth, year)

model_resample_depth_average <- model_resample_depth %>%
  group_by(depth) %>%
  summarise(n = mean(n))

model_resample_depth_average %>%
  arrange(depth) %>%
  ggplot(aes(n, depth)) +
  geom_point(data = model_resample_depth,
             aes(n, depth, col = "All years")) +
  geom_point(aes(n, depth, col = "Average")) +
  geom_path(aes(n, depth, col = "Average")) +
  scale_color_brewer(palette = "Set1",
                     name = "",
                     direction = -1) +
  scale_y_reverse() +
  coord_cartesian(xlim = c(0,500),
                  ylim = c(6000,0)) +
  labs(title = "Depth distribution of random model subset")


# Calculate and plot depth distribution of GLODAP data
# Depths gridded to model depth levels for comparison
GLODAP_depth <- GLODAP %>%
  mutate(depth = as.numeric(as.character(cut(depth,
                     c(0,unique(model_resample_depth_average$depth)),
                     labels = unique(model_resample_depth_average$depth))))) %>% 
  count(depth, year)

GLODAP_depth_average <- GLODAP_depth %>%
  group_by(depth) %>%
  summarise(n = mean(n))

GLODAP_depth_average %>%
  arrange(depth) %>%
  ggplot(aes(n, depth)) +
  geom_point(data = GLODAP_depth,
             aes(n, depth, col = "All years")) +
  geom_point(aes(n, depth, col = "Average")) +
  geom_path(aes(n, depth, col = "Average")) +
  scale_color_brewer(palette = "Set1",
                     name = "",
                     direction = -1) +
  scale_y_reverse() +
  coord_cartesian(xlim = c(0,500),
                  ylim = c(6000,0)) +
  labs(title = "Depth distribution of GLODAP observations")

```

