---
title: "eMLR model loop"
author: "Donghe Zhu and Dr Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```


```{r library, message = FALSE, warning = FALSE}
library(tidyverse)
library(patchwork)
library(stars)
library(collapse)
library(lubridate)
library(rqdatatable)
library(gsw)
library(marelac)
```

```{r read_parameters, include = FALSE}
parameters <-
  read_rds(here::here("data",
                       "parameters.rds"))
```

```{r read_mask_files, include = FALSE}

basinmask <- read_csv(
  here::here(
    "data/World_Ocean_Atlas_2018/_summarized_files",
    "basin_mask_WOA18_AIP.csv"
  )
)
basinmask <- basinmask %>%
  select(-basin)

landmask <-  read_csv(
  here::here(
    "data/World_Ocean_Atlas_2018/_summarized_files",
    "land_mask_WOA18.csv"
  )
)

```

```{r read_section_global_coordinates, include = FALSE}

section_global_coordinates <- read_csv(
  here::here(
    "data/World_Ocean_Atlas_2018/_summarized_files",
    "section_global_coordinates.csv"
  )
)

```

```{r read_functions, include = FALSE}
source(here::here("code", "plotting_functions.R"))
```

```{r ggplot_theme, include = FALSE}
# ggplot theme for all plots
theme_set(theme_bw())
```


# Read GLODAP clean data and apply the basinmask

```{r read_GLODAP_clean_data}

GLODAP <- read_csv(here::here("data/GLODAPv2_2020",
                              "GLODAPv2.2020_Merged_clean.csv"))
GLODAP <- inner_join(GLODAP, basinmask) %>%
  select(-basin_AIP)

```

# Ramdomly subset cmorized model data

Here we ramdomly subset cmorized (1x1) data from the model with variable forcing, according to the number of GLODAP observations (latxlon) in each year from a previously cleaned file.

Besides, Model results are given in [mol m^-3^], whereas GLODAP data are in [µmol kg^-1^]. This refers to the variables:

- DIC  
- ALK  
- O2  
- NO3  
- PO4  
- SiO3  
- AOU (calculated)

For comparison, model results were converted from [mol m^-3^] to [µmol kg^-1^]

```{r set_absolute_path}

# set path to model output files
path_model_cmorized <-
  "/net/kryo/work/loher/CESM_output/RECCAP2/cmorized_Dec2020/split_monthly/"

# set path to model analysis/subsetting output files
path_model_analysis <-
  "/net/kryo/work/loher/CESM_output/RECCAP2/analysis/"

```


```{r random_sample_cmorized_variable_forcing_data, eval=FALSE}

# read in a random model
model <-
  read_ncdf(
    "/net/kryo/work/loher/CESM_output/RECCAP2/cmorized_Dec2020/split_monthly/dissic_CESM-ETHZ_A_1_gr_1982.nc"
  ) %>%
  as_tibble() %>%
  drop_na()

# convert longitudes and mutate month
model <- model %>%
  mutate(lon = if_else(lon < 20, lon + 360, lon)) %>%
  mutate(month = month(time_mon))

# only consider model grids within basinmask
model <- inner_join(model, basinmask)

# model grid with depth
model_grid_depth <- model %>%
  select(month, lat, lon, depth)

rm(model)

# set name of model to be subsetted
model_ID <- "A"

# for loop across years
years <- c("1982":"2019")

# set equal number of random model sampling will be made for each year
n = floor(nrow(GLODAP) / length(years))

for (i_year in years) {
  # for manual testing
 # i_year <- years[1]
  
  # random sample n = GLODAP / years from model grid depth
  model_resample_grid_depth <- sample_n(model_grid_depth, n) %>%
    arrange(lat, lon, depth, month) %>%
    mutate(year = i_year)
  
  # for loop across variables
  variables <-
    c("so", "thetao", "dissic", "talk", "o2", "no3", "po4", "si")
  # "SALT", "TEMP", "DIC", "ALK", "O2", "NO3", "PO4", "SiO3"
  
  for (i_variable in variables) {
    # Uncomment line below for manual testing
    #i_variable <- variables[1]
    
    # read list of all files
    file <-
      list.files(
        path = path_model_cmorized,
        pattern = paste(
          "^",
          i_variable,
          "_CESM-ETHZ_",
          model_ID,
          "_1_gr_",
          i_year,
          ".nc",
          sep = ""
        )
      )
    print(file)
    
    # read in data
    variable_data <-
      read_ncdf(paste(path_model_cmorized,
                      file,
                      sep = ""))
    
    # convert to tibble
    variable_data_tibble <- variable_data %>%
      as_tibble()
    
    # remove open link to nc file
    rm(variable_data)
    
    # remove na values -> do this early, to speed up subsequent tasks
    variable_data_tibble <-
      variable_data_tibble %>%
      filter(!is.na(!!sym(i_variable)))
    
    # convert longitudes
    variable_data_tibble <- variable_data_tibble %>%
      mutate(lon = if_else(lon < 20, lon + 360, lon))
    
    # only consider model grids within basinmask
    variable_data_tibble <- inner_join(variable_data_tibble, basinmask) %>%
      select(-basin_AIP)
    
    # mutate variables
    variable_data_tibble <- variable_data_tibble %>%
      mutate(month = month(time_mon), !!sym(i_variable) := as.numeric(!!sym(i_variable))) %>%
      select(-time_mon)
    
    # random sample for each model variable in specific year
    model_resample_grid_depth <-
      left_join(model_resample_grid_depth, variable_data_tibble)
  }
  
  # add random sample model subset for each year together
  if (exists("model_resample")) {
    model_resample <-
      bind_rows(model_resample, model_resample_grid_depth)
  }
  
  if (!exists("model_resample")) {
    model_resample <- model_resample_grid_depth
  }
  
}

# unit transfer from mol/m3 to µmol/kg
model_resample <- model_resample %>%
  mutate(
    rho = gsw_pot_rho_t_exact(
      SA = so,
      t = thetao,
      p = depth,
      p_ref = 10.1325
    ),
    dissic = dissic * (1000000 / rho),
    talk = talk * (1000000 / rho),
    o2 = o2 * (1000000 / rho),
    no3 = no3 * (1000000 / rho),
    po4 = po4 * (1000000 / rho),
    si = si * (1000000 / rho)
  )

# calculate AOU
model_resample <- model_resample %>%
  mutate(
    oxygen_sat_m3 = gas_satconc(
      S = so,
      t = thetao,
      P = 1.013253,
      species = "O2"
    ),
    oxygen_sat_kg = oxygen_sat_m3 * (1000 / rho),
    aou = oxygen_sat_kg - o2
  ) %>%
  select(-oxygen_sat_kg,-oxygen_sat_m3)

# write file for random model sampling
model_resample %>%
  write_csv(paste(
    paste(path_model_analysis, "glodapv2_cmorized_synthetic/", sep = ""),
    "random_cmorized_subset_w_aou.csv",
    sep = ""
  ))

```


```{r plot_random_sample_cmorized_data_grid}

# read in random model sampling file
model_resample <-
  read_csv(paste(
    paste(path_model_analysis, "glodapv2_cmorized_synthetic/", sep = ""),
    "random_cmorized_subset_w_aou.csv",
    sep = ""
  ))

# plot random sampling cmorized grids in each year
years <- c("1982", "1990", "2000", "2010", "2019")
for (i_year in years) {
  model_resample_year <- model_resample %>%
    filter(year == i_year)
  
  print(
    ggplot() +
      geom_point(data = model_resample_year,
                 aes(lon, lat), colour = "red") +
      labs(
        title = paste("Random Model Sampling of year", i_year),
        x = "Longitude",
        y = "Latitude"
      ) +
      geom_raster(data = landmask,
                  aes(lon, lat), fill = "grey80")
  )
}

```

