---
title: "eMLR model loop"
author: "Donghe Zhu and Dr Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```


```{r library, message = FALSE, warning = FALSE}
library(tidyverse)
library(patchwork)
library(stars)
library(collapse)
library(lubridate)
library(rqdatatable)
```

```{r read_parameters, include = FALSE}
parameters <-
  read_rds(here::here("data",
                       "parameters.rds"))
```

```{r read_mask_files, include = FALSE}

basinmask <- read_csv(
  here::here(
    "data/World_Ocean_Atlas_2018/_summarized_files",
    "basin_mask_WOA18_AIP.csv"
  )
)
basinmask <- basinmask %>%
  select(-basin)

landmask <-  read_csv(
  here::here(
    "data/World_Ocean_Atlas_2018/_summarized_files",
    "land_mask_WOA18.csv"
  )
)

```

```{r read_section_global_coordinates, include = FALSE}

section_global_coordinates <- read_csv(
  here::here(
    "data/World_Ocean_Atlas_2018/_summarized_files",
    "section_global_coordinates.csv"
  )
)

```

```{r read_functions, include = FALSE}
source(here::here("code", "plotting_functions.R"))
```

```{r ggplot_theme, include = FALSE}
# ggplot theme for all plots
theme_set(theme_bw())
```


# Read GLODAP clean data

```{r read_GLODAP_clean_data}

GLODAP <- read_csv(here::here("data/GLODAPv2_2020",
                              "GLODAPv2.2020_Merged_clean.csv"))

```

# Subset model data according to GLODAP observation

Here we subset regridded (1x1) data from the model with variable forcing, according to the presence of GLODAP observations in a previously cleaned file.

Simultanously, we calculate annual model summary statistics and annual average DIC fields.

```{r set_absolute_path}

# set path to model output files
path_model_regridded <-
  "/net/kryo/work/loher/CESM_output/RECCAP2/regridded/"

# set path to model analysis/subsetting output files
path_model_analysis <-
  "/nfs/kryo/work/loher/CESM_output/RECCAP2/analysis/"

```



```{r subset_regridded_variable_forcing_data, eval=FALSE}


# set name of modelled to be subsetted
model_ID <- "GIAF_JRA_2020"

# for loop across years
years <- c("1982":"2018")

for (i_year in years) {
  # for manual testing
  # i_year <- years[1]
  
  # select GLODAP data for that year
  GLODAP_year <- GLODAP %>%
    filter(year == i_year)
  
  # for loop across variables
  variables <-
    c("DIC")
  # variables <-
  #   c("DIC", "ALK", "O2", "NO3", "PO4", "SiO3", "SALT", "TEMP")
  
  # create month x lat x lon grid of observations
  grid_subset <- GLODAP_year %>%
    select(month, lat, lon) %>%
    unique()
  
  # create month x lat x lon x depth grid of observations
  obs_depth <- GLODAP_year %>%
    select(month, lat, lon, depth) %>%
    unique()
  
  for (i_variable in variables) {
    # Uncomment line below for manual testing
    # i_variable <- variables[1]
    
    # read list of all files
    file <-
      list.files(
        path = path_model_regridded,
        pattern = paste(model_ID, "_", i_variable, "_", i_year, ".nc", sep = "")
      )
    print(file)
    
    # read in data
    variable_data <-
      read_ncdf(paste(path_model_regridded,
                      file,
                      sep = ""))
    
    # convert to tibble
    variable_data_tibble <- variable_data %>%
      as_tibble()
    
    # remove open link to nc file
    rm(variable_data)
    
    # remove na values -> do this early, to speed up subsequent tasks
    variable_data_tibble <-
      variable_data_tibble %>%
      filter(!is.na(!!sym(i_variable)))
    
    # convert longitudes
    variable_data_tibble <- variable_data_tibble %>%
      mutate(lon = if_else(lon < 20, lon + 360, lon))
    
    # mutate variables and filter NaN value
    variable_data_tibble <- variable_data_tibble %>%
      rename(depth = z_t) %>%
      mutate(
        depth = depth / 100,
        month = month(time),!!sym(i_variable) := as.numeric(!!sym(i_variable))
      ) %>%
      select(-time)
    
    # calculate model summary stats
    stats <- variable_data_tibble %>%
      pull(!!sym(i_variable)) %>%
      summary()
    
    stats <- c(year = i_year, variable = i_variable, stats)
    
    if (exists("stats_summary")) {
      stats_summary <- bind_rows(stats_summary, stats)
    }
    
    if (!exists("stats_summary")) {
      stats_summary <- stats
    }
    
    rm(stats)
    
    # subset model at month x lat x lon grid of observations
    model_subset <-
      inner_join(grid_subset, variable_data_tibble)
    
    # join model and month x lat x lon x depth grid of observations
    model_obs <- full_join(model_subset, obs_depth)
    
    # calculate nr of observations per month x lat x lon grid
    model_obs <- model_obs %>%
      group_by(month, lat, lon) %>%
      mutate(n = sum(!is.na(!!sym(i_variable)))) %>%
      ungroup()
    
    # interpolate variable value from model to observation depth
    model_obs_interpo <- model_obs %>%
      filter(n > 1) %>%
      group_by(lon, lat, month) %>%
      arrange(depth) %>%
      mutate(!!sym(i_variable) := approxfun(depth,!!sym(i_variable), rule = 2)(depth)) %>%
      ungroup()
    
    # subsetted interpolated values at observation depth
    model_obs_interpo <-
      inner_join(obs_depth, model_obs_interpo) %>%
      select(-n) %>%
      mutate(year = as.numeric(i_year))
    
    # rename interpolated model variable to indicate as model output
    model_obs_interpo <- model_obs_interpo %>%
      rename(!!sym(paste(i_variable, "model", sep = "_")) := !!sym(i_variable))
    
    # add model subset to GLODAP
    GLODAP <-
      natural_join(
        GLODAP,
        model_obs_interpo,
        by = c("year", "month", "lat", "lon", "depth"),
        jointype = "FULL"
      )
    
    # calculate annual average variable
    variable_data_tibble_annual_average <- variable_data_tibble %>%
      fselect(-month) %>%
      fgroup_by(lat, lon, depth) %>% {
        add_vars(fgroup_vars(., "unique"),
                 fmean(., keep.group_vars = FALSE))
      }
    
    # select surface annual average variable
    variable_data_tibble_annual_average_surface <-
      variable_data_tibble_annual_average %>%
      filter(depth == min(depth))
    
    # surface map of variable
    surface_GLODAP(
      variable_data_tibble_annual_average_surface,
      obs_depth,
      model_obs_interpo,
      i_variable
    )
    
    ggsave(
      here::here(
        "output/model_subsetting/surface_model_GLODAP_obs",
        paste(i_variable, "_", i_year, ".png", sep = "")
      ),
      width = 5,
      height = 3
    )
    
    # GLODAP with model subset:red point
    # GLODAP without model subset:black point
    
    # combine annual average model file together
    if (i_variable == "DIC") {
      # write file for annually averaged model field
      variable_data_tibble_annual_average %>%
        write_csv(paste(
          paste(path_model_analysis, "climatologies/", model_ID, sep = ""),
          paste("/DIC_annual_average_", i_year, ".csv", sep = ""), sep = ""
        ))
      
      
    }
    
  }
  
}

# write file for GLODAP observations + subsetted model variables
GLODAP %>%
  write_csv(paste(
    paste(path_model_analysis, "glodapv2_2020_synthetic_data/", sep = ""),
    "GLODAP_with_model_subset.csv",
    sep = ""
  ))

# write file for model summary statistics
stats_summary %>%
  write_csv(paste(
    paste(path_model_analysis, "diagnostics/", sep = ""),
    paste(model_ID, "summary_stats.csv", sep = "_")
  ))

```

# Test for interpolation

```{r test_for_interpolation, eval=FALSE}

# subset model at month x lat x lon grid of observations
model_subset <-
  inner_join(grid_subset, variable_data_tibble)

# join model and month x lat x lon x depth grid of observations
model_obs <- full_join(model_subset, obs_depth)

# calculate nr of observations per month x lat x lon grid
model_obs <- model_obs %>%
  group_by(month, lat, lon) %>%
  mutate(n = sum(!is.na(!!sym(i_variable)))) %>%
  ungroup()

# interpolate variable value from model to observation depth
model_obs_interpo <- model_obs %>%
  filter(n > 1) %>%
  group_by(lon, lat, month) %>%
  arrange(month, lat, lon, depth) %>%
  mutate(!!sym(paste(i_variable, "interpolate", sep = "_")) := approxfun(depth,!!sym(i_variable), rule = 2)(depth)) %>%
  ungroup()

ggplot() +
  geom_path(
    data = model_obs_interpo %>%
      filter(lat == 1.5, lon == 316.5, !is.na(!!sym(i_variable)), month == 1) %>%
      arrange(depth),
    aes(!!sym(i_variable), depth, col = "model")
  ) +
  geom_point(
    data = model_obs_interpo %>%
      filter(lat == 1.5, lon == 316.5, month == 1),
    aes(!!sym(
      paste(i_variable, "interpolate", sep = "_")
    ), depth, col = "interpolated")
  ) +
  geom_point(
    data = model_obs_interpo %>%
      filter(lat == 1.5, lon == 316.5, !is.na(!!sym(i_variable)), month == 1) %>%
      arrange(depth),
    aes(!!sym(i_variable), depth, col = "model")
  ) +
  scale_y_reverse() +
  scale_color_brewer(palette = "Dark2", name = "") +
  labs(title = paste("Interpolation to sampling depth -", i_variable, "of year", i_year))

```

# Annual average DIC fields

To derive Cant estimates for certain years, we calculate annual average DIC fields also for the output of a model with normal forcing.

```{r calculate_normal_forcing_annual_average_DIC, eval=FALSE}

# set path to model output files
path_model_regridded <-
  "/net/kryo/work/loher/CESM_output/RECCAP2/regridded/"

# set name of modelled to be subsetted
model_ID <- "GECO_NYF_cstCO2"

# for loop across years
years <- c("1982":"2018")
# years <- c("1982":"1983")

for (i_year in years) {
  # for manual testing
  # i_year <- years[1]
  
  # read list of all files
  file <-
    list.files(path = path_model_regridded,
               pattern = paste(model_ID, "_DIC_", i_year, ".nc", sep = ""))
  print(file)
  
  # read in data
  normal_data <-
    read_ncdf(paste(path_model_regridded,
                    file,
                    sep = ""))
  
  # convert to tibble
  normal_data_tibble <- normal_data %>%
    as_tibble()
  
  # remove open link to nc file
  rm(normal_data)
  
  # remove na values -> do this early, to speed up subsequent tasks
  normal_data_tibble <-
    normal_data_tibble %>%
    filter(!is.na(DIC))
  
  # convert longitudes
  normal_data_tibble <- normal_data_tibble %>%
    mutate(lon = if_else(lon < 20, lon + 360, lon))
  
  # mutate variables and filter NaN value
  normal_data_tibble <- normal_data_tibble %>%
    rename(depth = z_t) %>%
    mutate(depth = depth / 100,
           month = month(time),
           DIC = as.numeric(DIC)) %>%
    select(-time)
  
  # calculate annual average DIC
  normal_data_tibble_annual_average <- normal_data_tibble %>%
    fselect(-month) %>%
    fgroup_by(lat, lon, depth) %>% {
      add_vars(fgroup_vars(., "unique"),
               fmean(., keep.group_vars = FALSE))
    }
  
  # write file for annually averaged model DIC
  normal_data_tibble_annual_average %>%
    write_csv(paste(
    paste(path_model_analysis, "climatologies/",model_ID, sep = ""),
      paste("/DIC_annual_average_", i_year, ".csv", sep = "")
    ))
  
}

```

# Calculate model Cant

The model Cant can be calculate as the difference between two model runs. Currently, this is performed for year 1994 and 2007.

```{r calculate_Cant, eval=FALSE}

# read in variable and normal forcing DIC data
years <- c("1994", "2007")

# GLODAP <- read_csv(here::here("data/GLODAPv2_2020",
#                               "GLODAPv2.2020_Merged_clean.csv"))

for (i_year in years) {
  # for manual testing
  # i_year <- "1995"
  
  # read in variable forcing annual average DIC data
  variable_annual_average_DIC <-
    read_csv(paste(
    paste(path_model_analysis, "climatologies/GIAF_JRA_2020/", sep = ""),
      paste("DIC_annual_average_", i_year, ".csv", sep = ""), sep = ""
    )) %>% 
    rename(DIC_variable = DIC)
  
  # read in normal forcing annual average DIC data
  normal_annual_average_DIC <-
    read_csv(paste(
    paste(path_model_analysis, "climatologies/GECO_NYF_cstCO2/", sep = ""),
      paste("DIC_annual_average_", i_year, ".csv", sep = ""), sep = ""
    )) %>% 
    rename(DIC_normal = DIC)

  # calculate Cant
  cant <-
    full_join(variable_annual_average_DIC, normal_annual_average_DIC) %>%
    mutate(Cant = DIC_variable - DIC_normal)
  
  rm(variable_annual_average_DIC, normal_annual_average_DIC)
  
  cant <- inner_join(basinmask, cant)
  
  # global surface Cant plot
  print(map_global(cant, "Cant"))
  
  # global section Cant plot
  print(section_global(cant, "Cant"))
  
  # zonal section Cant plot
  for (i_basin_AIP in unique(cant$basin_AIP)) {
    breaks <- c(seq(-30, 70, 10), Inf)
    print(section_zonal(cant,
                        i_basin_AIP = i_basin_AIP,
                        "Cant"))
  }
}

```

